---
title: "MGSC661 Final Project Code"
output: html_notebook
---
```{r}
library(randomForest)
library(dplyr)
library(earth)
library(party)
library(e1071)
library(reshape2)
library(corrplot)
library(ggplot2)
library(Hmisc)
library(ggpubr)
library(stargazer)
library(Boruta)
library(car)
library(gbm)
require(caTools)
library(ggfortify)
library(lmtest)
library(caret)
library(pROC)
library(stargazer)
library(InformationValue)

```

# 1. Data preprocessing

```{r}
Protein = Protein_Supply_Quantity_Data
attach(Protein)
```

## 1.1 check missing values
```{r}
colSums(is.na(Protein))
```

```{r}
install.packages("VIM")
library(VIM)
```

## 1.2 visualize missing values
```{r}
pattern <- aggr(Protein, col=c('chartreuse3','violetred2'),
                    numbers=TRUE, sortVars=TRUE,
                    labels=names(Protein), cex.axis=.8,
                    gap=2, ylab=c("Missing data","Pattern"))
```


## 1.3 impute missing values
```{r}
install.packages("mice")
library(mice)
```

```{r}
imputed_Data <- mice(Protein, m=5, maxit = 50, method = 'cart', seed = 500)
```

```{r}
# This the cleaned dataset
clean_data <- complete(imputed_Data,2)
```

```{r}
attach(clean_data)
```

```{r}
#write.csv(clean_data,"clean_data.csv", row.names = FALSE)
```


## 1.4 Finalize Covid Recovery datasets
```{r}
Protein_Recovered = clean_data[,c(2:24,29)]
```


# 2. Feature Selection

## 2.1 Features Selection using Bortua


```{r}
# for Recovered
`%notin%` <- Negate(`%in%`)
boruta_output <- Boruta(Recovered ~ ., data=Protein_Recovered, doTrace=2)
boruta_signif <- names(boruta_output$finalDecision[boruta_output$finalDecision %in% c("Confirmed", "Tentative")])
boruta_not_signif <- names(boruta_output$finalDecision[boruta_output$finalDecision %notin% c("Confirmed", "Tentative")])
print(boruta_signif)
print(boruta_not_signif)
plot(boruta_output, cex.axis=.7, las=2, xlab="", main="Variable Importance for Recovered Rate") 

```
 
 
## 2.2 Features Selection using Random Forest

```{r}
# for Recovered
set.seed(0)
Recovered_Forest=randomForest(Recovered ~. #put all the x predictors here
                        , ntree=500, data=Protein_Recovered, importance=TRUE, na.action = na.omit)
varImpPlot(Recovered_Forest)

```
```{r}
round(importance(Recovered_forest),2)
```
 
 
 
## 2.4 Select predictors

```{r}
colnames(Protein_Recovered)
```


```{r}
Protein_Recovered_selected = subset(Protein_Recovered,select=c(Recovered,Animal.Products,Vegetable.Oils,Animal.fats,Vegetal.Products,Fish..Seafood,Milk...Excluding.Butter,Oilcrops,Eggs))
```

```{r}
summary()
```

```{r}
colnames(Protein_Recovered_selected)
```


```{r}
# summary statistics for selected features
print(summary(Protein_Recovered_selected$Animal.Products))
print(summary(Protein_Recovered_selected$Vegetable.Oils))
print(summary(Protein_Recovered_selected$Animal.fats))
print(summary(Protein_Recovered_selected$Vegetal.Products))
print(summary(Protein_Recovered_selected$Fish..Seafood))
print(summary(Protein_Recovered_selected$Milk...Excluding.Butter))
print(summary(Protein_Recovered_selected$Oilcrops))
print(summary(Protein_Recovered_selected$Eggs))
```


# 3. Data Exploration and Description

## 3.1 Data Visualization

### Histograms

```{r}
# for Recovered
# for (i in colnames(Protein_Recovered_selected)){
#   boxplot(Protein_Recovered_selected[[i]],col="cadetblue3",main=paste("Boxplot of", names(Protein_Recovered_selected[i])),xlab = paste(names(Protein_Recovered_selected[i])))
#   hist(Protein_Recovered_selected[[i]],col="cadetblue3",main=paste("Histogram of", names(Protein_Recovered_selected[i])),xlab = paste(names(Protein_Recovered_selected[i])))
# }

```


```{r}
# Matrix of histograms for features
colnames(Protein_Recovered_selected)
gg1=ggplot(Protein_Recovered_selected, aes(x = Animal.Products),fill='Black') + geom_histogram(color = "black", fill='dodgerblue3',binwidth=2.5)+
  labs( x = "Animal Products", y = "Frequency")+
  theme_bw()+scale_fill_manual(name = "Consumption Level", values = c("#92C8EB")) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5), legend.text = element_text(size = 12), axis.title = element_text(size = 12, face = "bold"))
gg1

gg2=ggplot(Protein_Recovered_selected, aes(x = Vegetable.Oils),fill='Black') + geom_histogram(color = "black", fill='dodgerblue3',binwidth=0.01)+
  labs( x = "Vegetable Oils", y = "Frequency")+
  theme_bw()+scale_fill_manual(name = "Consumption Level", values = c("#92C8EB")) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5), legend.text = element_text(size = 12), axis.title = element_text(size = 12, face = "bold"))
gg2

gg3=ggplot(Protein_Recovered_selected, aes(x = Animal.fats),fill='Black') + geom_histogram(color = "black", fill='dodgerblue3',binwidth=0.05)+
  labs( x = "Animal fats", y = "Frequency")+
  theme_bw()+scale_fill_manual(name = "Consumption Level", values = c("#92C8EB")) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5), legend.text = element_text(size = 12), axis.title = element_text(size = 12, face = "bold"))
gg3

gg4=ggplot(Protein_Recovered_selected, aes(x = Vegetal.Products),fill='Black') + geom_histogram(color = "black", fill='dodgerblue3',binwidth=2.5)+
  labs( x = "Vegetal Products", y = "Frequency")+
  theme_bw()+scale_fill_manual(name = "Consumption Level", values = c("#92C8EB")) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5), legend.text = element_text(size = 12), axis.title = element_text(size = 12, face = "bold"))
gg4  

gg5=ggplot(Protein_Recovered_selected, aes(x = Fish..Seafood),fill='Black') + geom_histogram(color = "black", fill='dodgerblue3',binwidth=1)+
  labs( x = "Seafood", y = "Frequency")+
  theme_bw()+scale_fill_manual(name = "Consumption Level", values = c("#92C8EB")) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5), legend.text = element_text(size = 12), axis.title = element_text(size = 12, face = "bold"))
gg5

gg6=ggplot(Protein_Recovered_selected, aes(x = Milk...Excluding.Butter),fill='Black') + geom_histogram(color = "black", fill='dodgerblue3',binwidth=1)+
  labs( x = "Milk - Excluding Butter", y = "Frequency")+
  theme_bw()+scale_fill_manual(name = "Consumption Level", values = c("#92C8EB")) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5), legend.text = element_text(size = 12), axis.title = element_text(size = 12, face = "bold"))
gg6

gg7=ggplot(Protein_Recovered_selected, aes(x = Oilcrops),fill='Black') + geom_histogram(color = "black", fill='dodgerblue3',binwidth=1)+
  labs( x = "Oilcrops", y = "Frequency")+
  theme_bw()+scale_fill_manual(name = "Consumption Level", values = c("#92C8EB")) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5), legend.text = element_text(size = 12), axis.title = element_text(size = 12, face = "bold"))
gg7

gg8=ggplot(Protein_Recovered_selected, aes(x = Eggs),fill='Black') + geom_histogram(color = "black", fill='dodgerblue3',binwidth=0.35)+
  labs( x = "Eggs", y = "Frequency")+
  theme_bw()+scale_fill_manual(name = "Consumption Level", values = c("#92C8EB")) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5), legend.text = element_text(size = 12), axis.title = element_text(size = 12, face = "bold"))
gg8

library(gridExtra) 

figure <- grid.arrange(gg1, gg2, gg3,gg4,gg5,gg6,gg7,gg8,
                    
                    ncol = 2, nrow = 4, top= textGrob('Distrubution of Each Data Column',gp=gpar(fontface="bold",fontsize=16))
                    )

figure
```
```{r}
# Histogram for Target variable
gg9=ggplot(Protein_Recovered_selected, aes(x = Recovered),fill='black') + geom_histogram(color = "Black",fill='dodgerblue3',binwidth=0.4)+
  labs( x = "Recovered Rate (% population)", y = "Frequency", title='Target Variable Distribution')+
  theme_bw()+scale_fill_manual(name = "Consumption Level", values = c("#92C8EB")) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5), legend.text = element_text(size = 12), axis.title = element_text(size = 12, face = "bold"))
gg9
```

### Density Plot
```{r}
gg1=ggplot(Protein_Recovered_selected, aes(x = Animal.Products),fill='Black') + geom_density(color = "black", fill='dodgerblue3')+
  labs( x = "Animal Products", y = "Density")+
  theme_bw()+scale_fill_manual(name = "Consumption Level", values = c("#92C8EB")) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5), legend.text = element_text(size = 12), axis.title = element_text(size = 12, face = "bold"))
gg1

gg2=ggplot(Protein_Recovered_selected, aes(x = Vegetable.Oils),fill='Black') + geom_density(color = "black", fill='dodgerblue3')+
  labs( x = "Vegetable Oils", y = "Density")+
  theme_bw()+scale_fill_manual(name = "Consumption Level", values = c("#92C8EB")) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5), legend.text = element_text(size = 12), axis.title = element_text(size = 12, face = "bold"))
gg2

gg3=ggplot(Protein_Recovered_selected, aes(x = Animal.fats),fill='Black') + geom_density(color = "black", fill='dodgerblue3')+
  labs( x = "Animal fats", y = "Density")+
  theme_bw()+scale_fill_manual(name = "Consumption Level", values = c("#92C8EB")) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5), legend.text = element_text(size = 12), axis.title = element_text(size = 12, face = "bold"))
gg3

gg4=ggplot(Protein_Recovered_selected, aes(x = Vegetal.Products),fill='Black') + geom_density(color = "black", fill='dodgerblue3')+
  labs( x = "Vegetal Products", y = "Density")+
  theme_bw()+scale_fill_manual(name = "Consumption Level", values = c("#92C8EB")) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5), legend.text = element_text(size = 12), axis.title = element_text(size = 12, face = "bold"))
gg4  

gg5=ggplot(Protein_Recovered_selected, aes(x = Fish..Seafood),fill='Black') + geom_density(color = "black", fill='dodgerblue3',)+
  labs( x = "Seafood", y = "Density")+
  theme_bw()+scale_fill_manual(name = "Consumption Level", values = c("#92C8EB")) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5), legend.text = element_text(size = 12), axis.title = element_text(size = 12, face = "bold"))
gg5

gg6=ggplot(Protein_Recovered_selected, aes(x = Milk...Excluding.Butter),fill='Black') + geom_density(color = "black", fill='dodgerblue3')+
  labs( x = "Milk - Excluding Butter", y = "Density")+
  theme_bw()+scale_fill_manual(name = "Consumption Level", values = c("#92C8EB")) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5), legend.text = element_text(size = 12), axis.title = element_text(size = 12, face = "bold"))
gg6

gg7=ggplot(Protein_Recovered_selected, aes(x = Oilcrops),fill='Black') + geom_density(color = "black", fill='dodgerblue3')+
  labs( x = "Oilcrops", y = "Density")+
  theme_bw()+scale_fill_manual(name = "Consumption Level", values = c("#92C8EB")) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5), legend.text = element_text(size = 12), axis.title = element_text(size = 12, face = "bold"))
gg7

gg8=ggplot(Protein_Recovered_selected, aes(x = Eggs),fill='Black') + geom_density(color = "black", fill='dodgerblue3')+
  labs( x = "Eggs", y = "Density")+
  theme_bw()+scale_fill_manual(name = "Consumption Level", values = c("#92C8EB")) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5), legend.text = element_text(size = 12), axis.title = element_text(size = 12, face = "bold"))
gg8

library(gridExtra) 

figure <- grid.arrange(gg1, gg2, gg3,gg4,gg5,gg6,gg7,gg8,
                    
                    ncol = 2, nrow = 4, top= textGrob('Distrubution of Each Data Column',gp=gpar(fontface="bold",fontsize=16))
                    )

figure
```
```{r}
# Density for Target variable
gg9=ggplot(Protein_Recovered_selected, aes(x = Recovered),fill='black') + geom_density(color = "Black",fill='dodgerblue3')+
  labs( x = "Recovered", y = "Density", title='Target Variable Distribution')+
  theme_bw()+scale_fill_manual(name = "Consumption Level", values = c("#92C8EB")) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5), legend.text = element_text(size = 12), axis.title = element_text(size = 12, face = "bold"))
gg9
```


### Box Plot
```{r}
gg1=ggplot(Protein_Recovered_selected, aes(x = Animal.Products),fill='Black') + geom_boxplot(color = "black", fill='dodgerblue3')+
  labs( x = "Animal Products", y = "")+
  theme_bw()+scale_fill_manual(name = "Consumption Level", values = c("#92C8EB")) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5), legend.text = element_text(size = 12), axis.title = element_text(size = 12, face = "bold"))
gg1

gg2=ggplot(Protein_Recovered_selected, aes(x = Vegetable.Oils),fill='Black') + geom_boxplot(color = "black", fill='dodgerblue3')+
  labs( x = "Vegetable Oils", y = "")+
  theme_bw()+scale_fill_manual(name = "Consumption Level", values = c("#92C8EB")) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5), legend.text = element_text(size = 12), axis.title = element_text(size = 12, face = "bold"))
gg2

gg3=ggplot(Protein_Recovered_selected, aes(x = Animal.fats),fill='Black') + geom_boxplot(color = "black", fill='dodgerblue3')+
  labs( x = "Animal fats", y = "")+
  theme_bw()+scale_fill_manual(name = "Consumption Level", values = c("#92C8EB")) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5), legend.text = element_text(size = 12), axis.title = element_text(size = 12, face = "bold"))
gg3

gg4=ggplot(Protein_Recovered_selected, aes(x = Vegetal.Products),fill='Black') + geom_boxplot(color = "black", fill='dodgerblue3')+
  labs( x = "Vegetal Products", y = "")+
  theme_bw()+scale_fill_manual(name = "Consumption Level", values = c("#92C8EB")) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5), legend.text = element_text(size = 12), axis.title = element_text(size = 12, face = "bold"))
gg4  

gg5=ggplot(Protein_Recovered_selected, aes(x = Fish..Seafood),fill='Black') + geom_boxplot(color = "black", fill='dodgerblue3',)+
  labs( x = "Seafood", y = "")+
  theme_bw()+scale_fill_manual(name = "Consumption Level", values = c("#92C8EB")) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5), legend.text = element_text(size = 12), axis.title = element_text(size = 12, face = "bold"))
gg5

gg6=ggplot(Protein_Recovered_selected, aes(x = Milk...Excluding.Butter),fill='Black') + geom_boxplot(color = "black", fill='dodgerblue3')+
  labs( x = "Milk - Excluding Butter", y = "")+
  theme_bw()+scale_fill_manual(name = "Consumption Level", values = c("#92C8EB")) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5), legend.text = element_text(size = 12), axis.title = element_text(size = 12, face = "bold"))
gg6

gg7=ggplot(Protein_Recovered_selected, aes(x = Oilcrops),fill='Black') + geom_boxplot(color = "black", fill='dodgerblue3')+
  labs( x = "Oilcrops", y = "")+
  theme_bw()+scale_fill_manual(name = "Consumption Level", values = c("#92C8EB")) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5), legend.text = element_text(size = 12), axis.title = element_text(size = 12, face = "bold"))
gg7

gg8=ggplot(Protein_Recovered_selected, aes(x = Eggs),fill='Black') + geom_boxplot(color = "black", fill='dodgerblue3')+
  labs( x = "Eggs", y = "")+
  theme_bw()+scale_fill_manual(name = "Consumption Level", values = c("#92C8EB")) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5), legend.text = element_text(size = 12), axis.title = element_text(size = 12, face = "bold"))
gg8

library(gridExtra) 

figure <- grid.arrange(gg1, gg2, gg3,gg4,gg5,gg6,gg7,gg8,
                    
                    ncol = 2, nrow = 4, top= textGrob('Distrubution of Each Data Column',gp=gpar(fontface="bold",fontsize=16))
                    )

figure
```


```{r}
# BoxPlot for Target variable
gg9=ggplot(Protein_Recovered_selected, aes(x = Recovered),fill='black') + geom_boxplot(color = "Black",fill='dodgerblue3')+
  labs( x = "Recovered", y = "", title='Target Variable Distribution')+
  theme_bw()+scale_fill_manual(name = "Consumption Level", values = c("#92C8EB")) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5), legend.text = element_text(size = 12), axis.title = element_text(size = 12, face = "bold"))
gg9
```




```{r}
for (i in colnames(Protein_Recovered_selected)){print(i)}
```


## 3.2 Skewness

```{r}
colnames(Protein_Recovered_selected)
```

```{r}
attach(Protein_Recovered_selected)
```


```{r}
#Skewness

skewness_data = data.frame(
  variables = colnames(Protein_Recovered_selected),
  skewness = round(c(skewness(Recovered),skewness(Animal.Products),skewness(Vegetable.Oils),skewness(Animal.fats),skewness(Vegetal.Products),skewness(Fish..Seafood),skewness(Milk...Excluding.Butter),skewness(Oilcrops),skewness(Eggs)),2)
)


skewness_data$type=ifelse(
  
  ((skewness_data$skewness<0.5) &  (skewness_data$skewness>-0.5)),"Symmetrical",
  ifelse(((skewness_data$skewness>1)|(skewness_data$skewness<=-1)),"Highly Skewed",
  "Moderately Skewed")
  
  )

                
print(skewness_data)

```

## 3.3 Colinearity Test (collinearity and correlation)

### 1st Way - Correlation Matrix and Heat map

```{r}
corr_matrix = cor(Protein_Recovered_selected[,-c(1)], method = c("pearson"))
round(corr_matrix,2)
```

```{r}
# heap map #1
corrplot(corr_matrix)
# melted_cormat <- melt(X.cor)
# head(melted_cormat)
```


```{r}

# Heap map #2
# melted_corr_matrix_Recovered <- melt(cor(Protein_Recovered_selected, method = c("pearson")))
# ggplot(data = melted_corr_matrix_Recovered,aes(Var1, Var2, fill = value))+
#   geom_tile(color = "white")+
#   scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
#                        midpoint = 0, limit = c(-1,1), space = "Lab", 
#                        name="Pearson\nCorrelation") +
#   theme_minimal()+ 
#   theme(axis.text.x = element_text(angle = 45, vjust = 1, 
#                                    size = 12, hjust = 1))+
#   coord_fixed()

```

### 2rd Way - VIF
```{r}
round(vif(lm(Recovered~.,data=Protein_Recovered_selected)),2)
```

```{r}
# remove collinear predictors
Protein_Recovered_nocoll = subset(Protein_Recovered_selected,select=-c(Vegetal.Products))
```

```{r}
round(vif(lm(Recovered~.,data=Protein_Recovered_nocoll)),2)
```


## 3.4 Outlier Test (Bonferroni Test)
```{r}
# For Recovered
outlierTest(lm(Recovered~., data=Protein_Recovered_nocoll))

```

```{r}
# Remove outliers
Protein_Recovered_nooutlier = Protein_Recovered_nocoll[-c(103),]
```

```{r}
Protein_Recovered_final = Protein_Recovered_nooutlier
```

```{r}
colnames(Protein_Recovered_final)
```

## 3.5 Data preparation for modelling

### Log-Transformation

```{r}
# apply log transformation to target variable
Protein_Recovered_trans = Protein_Recovered_final
Protein_Recovered_trans$Recovered = log(Protein_Recovered_final$Recovered+1)
# Protein_Recovered_trans$Vegetable.Oils = log(Protein_Recovered_final$Vegetable.Oils+1)
# Protein_Recovered_trans$Animal.fats = log(Protein_Recovered_final$Animal.fats+1)
# Protein_Recovered_trans$Fish..Seafood = log(Protein_Recovered_final$Fish..Seafood+1)
# Protein_Recovered_trans$Oilcrops = log(Protein_Recovered_final$Oilcrops+1)

```

```{r}
# re-check skewness
for (i in colnames(Protein_Recovered_trans)){
  boxplot(Protein_Recovered_trans[[i]],col="cadetblue3",main=paste("Boxplot of", names(Protein_Recovered_trans[i])),xlab = paste(names(Protein_Recovered_trans[i])))
  hist(Protein_Recovered_trans[[i]],col="cadetblue3",main=paste("Histogram of", names(Protein_Recovered_trans[i])),xlab = paste(names(Protein_Recovered_trans[i])))
}
```


### Train-test Split

```{r}
require(caTools)
set.seed(0)
sample = sample.split(Protein_Recovered_trans$Recovered, SplitRatio = .8)
Protein_train = subset(Protein_Recovered_trans, sample == TRUE)
Protein_test  = subset(Protein_Recovered_trans, sample == FALSE)
```



# 4. Regression Models

## 4.1 Decision Tree

```{r}
install.packages("tree")
library(tree)
install.packages("rpart.plot")
library(rpart.plot)
```

### Decision Tree for regression

```{r}
# Start with a overfitted tree
set.seed(0)
myoverfittedtree=rpart(Recovered~., data=Protein_train, control=rpart.control(cp=0.0000001))
plotcp(myoverfittedtree)
```

```{r}
# Find the otimal CP
opt_cp=myoverfittedtree$cptable[which.min(myoverfittedtree$cptable[,"xerror"]),"CP"]
opt_cp
```


```{r}
# Grow the optimal tree
set.seed(0)
mybesttree=rpart(Recovered~., data=Protein_train,control=rpart.control(cp=opt_cp))
summary(mybesttree)
```

```{r}
# visualize the tree
rpart.plot(mybesttree)
```

```{r}
# Calculate MSE (with inverse transformation)
predicted_Recovered=predict(mybesttree, newdata=Protein_test)
mean(((exp(predicted_Recovered)-1) - (exp(Protein_test$Recovered)-1))^2)
```

### Decision Tree for data interpretation (built with the whole dataset)

```{r}
# Start with a overfitted tree
set.seed(0)
myoverfittedtree2=rpart(Recovered~., data=Protein_Recovered_trans, control=rpart.control(cp=0.0000001))
plotcp(myoverfittedtree2)
```

```{r}
# Find the otimal CP
opt_cp2=myoverfittedtree2$cptable[which.min(myoverfittedtree2$cptable[,"xerror"]),"CP"]
opt_cp2
```

```{r}
# Grow the optimal tree
set.seed(0)
mybesttree2=rpart(Recovered~., data=Protein_Recovered_trans,control=rpart.control(cp=opt_cp2))
summary(mybesttree2)
```

```{r}
# visualize the tree
rpart.plot(mybesttree2)
```


## 4.2 Random Forest

### Grow a random forest and OOB test
```{r}
set.seed(0)
myforest=randomForest(Recovered~., data=Protein_train,ntree=440,importance = TRUE,na.action = na.omit)
myforest
```

### Calculate MSE (with inverse transformation)

```{r}
predicted_Recovered=predict(myforest, newdata=Protein_test)
mean(((exp(predicted_Recovered)-1) - (exp(Protein_test$Recovered)-1))^2)
```

## 4.3 Gradient Boosting

### Build a boosted forest
```{r}
set.seed (0)
boosted=gbm(Recovered~., data=Protein_train,distribution= "gaussian",n.trees=430, interaction.depth=3)
summary(boosted)
```

### Calculate MSE
```{r}
predicted_Recovered=predict(boosted, newdata=Protein_test, n.trees=430)
mean(((exp(predicted_Recovered)-1) - (exp(Protein_test$Recovered)-1))^2)
```


# 5. Principal Component Analysis

## 5.1 Categorize Recovery Rate

```{r}
summary(Protein_Recovered_final$Recovered)
```


```{r}
Protein_Recovered_final = Protein_Recovered_final %>% mutate(Recovered_Category =
                     case_when(Recovered <= 0.5 ~ "Low Recovery Rate", 
                               Recovered <= 3 ~ "Medium Recovery Rate",
                               Recovered > 3  ~ "High Recovery Rate")
)
```


## 5.2 Apply PCA

```{r}
pca=prcomp(Protein_Recovered_final[,-c(1,9)], scale=TRUE)
pca
```

```{r}
pcaplot = autoplot(pca, data = Protein_Recovered_final[,-c(1,9)], loadings = TRUE, loadings.label = TRUE, col=ifelse(Protein_Recovered_final$Recovered_Category == "Low Recovery Rate", "red", ifelse(Protein_Recovered_final$Recovered_Category == "High Recovery Rate","green3", "blue")))
pcaplot
```

```{r}
pve=(pca$sdev^2)/sum(pca$sdev^2)
par(mfrow=c(1,2))
plot(pve, ylim=c(0,1))
plot(cumsum(pve), ylim=c(0,1))
```




# 6. Extension

```{r}
install.packages('countrycode')
library(countrycode)
```

```{r}
Protein_Recovered_extension = clean_data[,c(1:24,29)]
```

```{r}
Protein_Recovered_extension$landmass <- countrycode(sourcevar = Protein_Recovered_extension[, "Country"],
                             origin = 'country.name',
                             destination = "continent")
glimpse(Protein_Recovered_extension)
```


```{r}
Protein_Recovered_agg = aggregate(Protein_Recovered_extension[,-c(1,26)], by=list(landmass=Protein_Recovered_extension$landmass), FUN=sum)
Protein_Recovered_agg
colnames(Protein_Recovered_agg)
```
```{r}
Protein_Recovered_agg_selected =subset(Protein_Recovered_agg,select=c(landmass,Recovered,Animal.Products,Vegetable.Oils,Animal.fats,Vegetal.Products,Fish..Seafood,Milk...Excluding.Butter,Oilcrops,Eggs))
Protein_Recovered_agg_selected
```

```{r}
colnames(Protein_Recovered_agg_selected)
```

```{r}
write.csv(Protein_Recovered_agg_selected,"aggregation.csv", row.names = FALSE)
```












